---
title: "Using joint analysis with multi-modal data: CITE-seq"
author: Lei Li <leil@uchicago.edu>
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
  html_document: default
  pdf_document: default
  knitrBootstrap::bootstrap_document:
    theme: amelia
    highlight: sunburst
    theme.chooser: TRUE
    highlight.chooser: TRUE
---


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# Load packages

```{r results='hide', results='hide', message=FALSE, warning=FALSE, warning=FALSE}
library(LinQView)
library(Seurat)
library(cowplot)
library(ggplot2)
```

# Step 1 Load data from 10X folder
Data can be downloaded from 10X website https://support.10xgenomics.com/single-cell-gene-expression/datasets/3.0.0/pbmc_10k_protein_v3
```{r results='hide', results='hide', message=FALSE, warning=FALSE, warning=FALSE}
# Load in the RNA UMI matrix
cbmc.data <- readDataFrom10X(dir = "../../../Data/10K/filtered_feature_bc_matrix/")
```

# Step 2 Create object
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- createObject(data = cbmc.data)
t2 <- Sys.time()
t2 - t1
```

# Step 3 Pre-process
###  1) Filter out unwanted cells (optional)
for this dataset, we don't need to filter out unwanted cells
```{r results='hide', message=FALSE, warning=FALSE}
cbmc <- subset(cbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < softThreshold(cbmc))
```

###  2) Remove unwanted genes (optional)
for this dataset, we don't need to filter out unwanted genes
```{r results='hide', message=FALSE, warning=FALSE}
# remove Ig genes
#cbmc <- removeGene(object = cbmc,pattern = '^IG[HKL]')
```

###  3) Normalization 
data Normalization for both ADT (CLR) and RNA (log)
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- dataNormalization(object = cbmc)
t2 <- Sys.time()
t2 - t1
```

###  4) Indentify HVGs for RNA data
Call seurat function to identify highly variable genes (HVG) for RNA data
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- FindVariableFeatures(object = cbmc)   # directly use Seurat Function
t2 <- Sys.time()
t2 - t1
```

###  5) Data scaling 
Scale data for both ADT and RNA
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- dataScaling(object = cbmc) 
t2 <- Sys.time()
t2 - t1
```


# Step 4 Linear dimension reduction (PCA)
directly call Seurat function for linear dimension reduction (PCA)
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- RunPCA(cbmc, features = VariableFeatures(object = cbmc), verbose = FALSE)   # directly use Seurat Function
t2 <- Sys.time()
t2 - t1
```

# Step 5 Determine number of PCs
call Seurat function JackStraw to determine number of PCs
```{r fig.width=6, fig.height=6, results='hide', message=FALSE, warning=FALSE}
#cbmc <- JackStraw(cbmc, num.replicate = 100)
#cbmc <- ScoreJackStraw(cbmc, dims = 1:20)
#JackStrawPlot(cbmc, dims = 1:15)
#ElbowPlot(cbmc)
```

# Step 6 Distance calculation and joint distance calculation
calculate cell-cell distances for RNA, ADT and joint. alpha was set to 0.5 as initial, number of PC was set to 20 by default.
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- jointDistance(object = cbmc, keep.rna = TRUE, keep.adt = TRUE)
t2 <- Sys.time()
t2 - t1
```
# Step 7 Non-linear dimension reduction (UMAP and t-SNE)
run UMAP as Non-linear dimension reduction for RNA, ADT and joint analysis. 
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- tsneFromDistane(object = cbmc, assay = "All")
t2 <- Sys.time()
t2 - t1
```
# Step 8 Clustering
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- clusteringFromDistance(object = cbmc, assay = "All", resolution = c(0.8,0.8,0.8))
t2 <- Sys.time()
t2 - t1
```

```{r echo=TRUE, dpi=300,fig.width=9, fig.height=7,results='hide', message=FALSE, warning=FALSE}
# contribution of two modalities
distHeatMap(object = cbmc)
```

# Step 9 Visualization ADT vs RNA vs Joint
### 1) Cell clusters
```{r , echo=TRUE, dpi=300,fig.width=9, fig.height=9, results='hide', message=FALSE, warning=FALSE}
#gridDimPlot(cbmc, wide.rel = 1.5, legend = FALSE, reduction.prefix = "tsne_", height.rel = 0.5)

plots <- generateGridDimPlot(cbmc, legend = FALSE, darkTheme = FALSE,cluster.lable.size = 8)
listPlot(object = plots, labels = "")

###### user also can only plot some of those plots by index, figure ident or figure map info
#listPlot(object = plots, fig.ident = "RNA")
#listPlot(object = plots, fig.ident = "RNA", fig.map = "RNA")

###### user can use plotInfo() function to get index, figure ident and figure map information, then plot figures by index
plotInfo(plots)
#listPlot(object = plots, fig.id = 1)
```

### 2) Heat maps
```{r  results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
# Heatmap for joint clusters
heatMapPlot(object = cbmc, group.by = "jointClusterID", height.rel = 3, adt.label = TRUE)
# Heatmap for RNA clusters
heatMapPlot(object = cbmc, group.by = "rnaClusterID", height.rel = 3, adt.label = TRUE)
# Heatmap for ADT clusters
heatMapPlot(object = cbmc, group.by = "adtClusterID", height.rel = 3, adt.label = TRUE)
```

```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=15, fig.height=3}
p1 <- VlnPlot(cbmc, features = "adt_CD4", pt.size = 0, group.by = 'jointClusterID') + NoLegend() 
p3 <- VlnPlot(cbmc, features = "adt_CD8a", pt.size = 0, group.by = 'jointClusterID') + NoLegend() 
p4 <- VlnPlot(cbmc, features = "adt_CD15", pt.size = 0, group.by = 'jointClusterID') + NoLegend()
p5 <- VlnPlot(cbmc, features = "adt_CD16", pt.size = 0, group.by = 'jointClusterID') + NoLegend() 
p6 <- VlnPlot(cbmc, features = "adt_CD127", pt.size = 0, group.by = 'jointClusterID') + NoLegend() 
plot_grid(p1,p3,p4,p5,p6,nrow = 1)
```


# gene difference
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=8, fig.height=7}
a1 <- FeatureScatter(cbmc, feature1 = "adt_CD16", feature2 = "rna_FCGR3A") + NoLegend() + theme(text=element_text(size=16, family="serif")) + LightTheme()

x <- cbmc@assays[["RNA"]]@counts["FCGR3A",]
cells <- which(x == 0)
y <- cbmc@assays[["ADT"]]@data["CD16",cells]
y <- as.data.frame(y)
p1 <- ggplot(y, aes(x=y)) + geom_histogram(fill = 'skyblue', color = 'grey30') + scale_y_log10() + ggtitle("FCGR3A gene negative cells") + LightTheme() + theme(text=element_text(size=16, family="serif"))

a2 <- FeaturePlot(cbmc, features = "adt_CD16", reduction = "tsne_joint") + NoLegend() + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()
a3 <- FeaturePlot(cbmc, features = "rna_FCGR3A", reduction = "tsne_joint") + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()

b1 <- FeatureScatter(cbmc, feature1 = "adt_CD19", feature2 = "rna_CD19") + NoLegend() + theme(text=element_text(size=16, family="serif")) + LightTheme()

x <- cbmc@assays[["RNA"]]@counts["CD19",]
cells <- which(x == 0)
y <- cbmc@assays[["ADT"]]@data["CD19",cells]
y <- as.data.frame(y)
p2 <- ggplot(y, aes(x=y)) + geom_histogram(fill = 'skyblue', color = 'grey30') + scale_y_log10() + ggtitle("CD19 gene negative cells") + LightTheme() + theme(text=element_text(size=16, family="serif"))

b2 <- FeaturePlot(cbmc, features = "adt_CD19", reduction = "tsne_joint") + NoLegend() + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()
b3 <- FeaturePlot(cbmc, features = "rna_CD19", reduction = "tsne_joint") + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()

c1 <- FeatureScatter(cbmc, feature1 = "adt_CD3", feature2 = "rna_CD3E") + NoLegend() + theme(text=element_text(size=16, family="serif")) + LightTheme()

x <- cbmc@assays[["RNA"]]@counts["CD3E",]
cells <- which(x == 0)
y <- cbmc@assays[["ADT"]]@data["CD3",cells]
y <- as.data.frame(y)
p3 <- ggplot(y, aes(x=y)) + geom_histogram(fill = 'skyblue', color = 'grey30') + scale_y_log10() + ggtitle("CD3E gene negative cells") + LightTheme() + theme(text=element_text(size=16, family="serif"))

c2 <- FeaturePlot(cbmc, features = "adt_CD3", reduction = "tsne_joint") + NoLegend() + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()
c3 <- FeaturePlot(cbmc, features = "rna_CD3E", reduction = "tsne_joint") + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()

d1 <- FeatureScatter(cbmc, feature1 = "adt_CD8a", feature2 = "rna_CD8A") + NoLegend() + theme(text=element_text(size=16, family="serif")) + LightTheme()

x <- cbmc@assays[["RNA"]]@counts["CD8A",]
cells <- which(x == 0)
y <- cbmc@assays[["ADT"]]@data["CD8a",cells]
y <- as.data.frame(y)
p4 <- ggplot(y, aes(x=y)) + geom_histogram(fill = 'skyblue', color = 'grey30') + scale_y_log10() + ggtitle("CD8A gene negative cells") + LightTheme() + theme(text=element_text(size=16, family="serif"))

d2 <- FeaturePlot(cbmc, features = "adt_CD8a", reduction = "tsne_joint") + NoLegend() + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()
d3 <- FeaturePlot(cbmc, features = "rna_CD8A", reduction = "tsne_joint") + theme(text=element_text(size=16, family="serif")) + xlab("t-SNE1") + ylab("t-SNE2") + LightTheme()

plot_grid(a1,p1,a2,a3,b1,p2,b2,b3,c1,p3,c2,c3,d1,p4,d2,d3, ncol = 4, rel_widths = c(1,1,1,1.2))
```

