---
title: "Using joint analysis with multi-modal data: CITE-seq"
author: Lei Li <leil@uchicago.edu>
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
  html_document: default
  pdf_document: default
  knitrBootstrap::bootstrap_document:
    theme: amelia
    highlight: sunburst
    theme.chooser: TRUE
    highlight.chooser: TRUE
---


```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.lazy = FALSE,
  tidy = TRUE
)
```

# Load packages

```{r results='hide', results='hide', message=FALSE, warning=FALSE, warning=FALSE}
library(LinQView)
library(Seurat)
library(cowplot)
library(ggplot2)
```

# CITE-seq dataset
In this tutorial, we use a public CITE-seq dataset to illustrate Joint analysis using LinQ-seq.
Data could be download from NCBI: 
RNA (ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE100nnn/GSE100866/suppl/GSE100866_CBMC_8K_13AB_10X-RNA_umi.csv.gz) 
ADT (ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE100nnn/GSE100866/suppl/GSE100866_CBMC_8K_13AB_10X-ADT_umi.csv.gz)

# Step 1 Load data from 10X folder
```{r results='hide', results='hide', message=FALSE, warning=FALSE, warning=FALSE}
# Load in the RNA UMI matrix
cbmc.rna <- as.sparse(x = read.csv(file = "../../../Data/citeseq/GSE100866_CBMC_8K_13AB_10X-RNA_umi.csv.gz", sep = ",", header = TRUE, row.names = 1))
cbmc.rna <- CollapseSpeciesExpressionMatrix(object = cbmc.rna)
# Load in the ADT UMI matrix
cbmc.adt <- as.sparse(x = read.csv(file = "../../../Data/citeseq/GSE100866_CBMC_8K_13AB_10X-ADT_umi.csv.gz", sep = ",", header = TRUE, row.names = 1))
cbmc.adt <- cbmc.adt[setdiff(x = rownames(x = cbmc.adt), y = c("CCR5", "CCR7", "CD10")), ]
```

# Step 2 Create object
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- createObject(rna = cbmc.rna, adt = cbmc.adt)
t2 <- Sys.time()
t2 - t1
```

# Step 3 Pre-process
###  1) Filter out unwanted cells (optional)
for this dataset, we don't need to filter out unwanted cells
```{r results='hide', message=FALSE, warning=FALSE}
cbmc <- subset(cbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < softThreshold(cbmc))
```

###  2) Remove unwanted genes (optional)
for this dataset, we don't need to filter out unwanted genes
```{r results='hide', message=FALSE, warning=FALSE}
# remove Ig genes
#cbmc <- removeGene(object = cbmc,pattern = '^IG[HKL]')
```

###  3) Normalization 
data Normalization for both ADT (CLR) and RNA (log)
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- dataNormalization(object = cbmc)
t2 <- Sys.time()
t2 - t1
```

###  4) Indentify HVGs for RNA data
Call seurat function to identify highly variable genes (HVG) for RNA data
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- FindVariableFeatures(object = cbmc)   # directly use Seurat Function
t2 <- Sys.time()
t2 - t1
```

###  5) Data scaling 
Scale data for both ADT and RNA
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- dataScaling(object = cbmc) 
t2 <- Sys.time()
t2 - t1
```


# Step 4 Linear dimension reduction (PCA)
directly call Seurat function for linear dimension reduction (PCA)
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- RunPCA(cbmc, features = VariableFeatures(object = cbmc), verbose = FALSE)   # directly use Seurat Function
t2 <- Sys.time()
t2 - t1
```

# Step 5 Determine number of PCs
call Seurat function JackStraw to determine number of PCs
```{r fig.width=6, fig.height=6, results='hide', message=FALSE, warning=FALSE}
#cbmc <- JackStraw(cbmc, num.replicate = 100)
#cbmc <- ScoreJackStraw(cbmc, dims = 1:20)
#JackStrawPlot(cbmc, dims = 1:20)
#ElbowPlot(cbmc)
```

# Step 6 Distance calculation and joint distance calculation
calculate cell-cell distances for RNA, ADT and joint. number of PC was set to 20 by default.
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- jointDistance(object = cbmc, keep.rna = TRUE, keep.adt = TRUE, dims = 25)
t2 <- Sys.time()
t2 - t1
```
# Step 7 Non-linear dimension reduction (UMAP and t-SNE)
run UMAP as Non-linear dimension reduction for RNA, ADT and joint analysis. 
```{r results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- tsneFromDistane(object = cbmc, assay = "All")
t2 <- Sys.time()
t2 - t1
```
# Step 8 Clustering
```{r echo=TRUE, dpi=300,fig.width=7, fig.height=5, results='hide', message=FALSE, warning=FALSE}
t1 <- Sys.time()
cbmc <- clusteringFromDistance(object = cbmc, assay = "All", resolution = c(0.9,0.9,0.9))
t2 <- Sys.time()
t2 - t1
# contribution of two modalities
distHeatMap(object = cbmc)
```

# Step 9 Visualization ADT vs RNA vs Joint
### 1) Cell clusters
```{r , echo=TRUE, dpi=300,fig.width=10, fig.height=10, results='hide', message=FALSE, warning=FALSE}
plots <- generateGridDimPlot(cbmc, legend = FALSE, darkTheme = FALSE)

listPlot(object = plots, align = "h")

###### user also can only plot some of those plots by index, figure ident or figure map info
#listPlot(object = plots, fig.ident = "RNA")
#listPlot(object = plots, fig.ident = "RNA", fig.map = "RNA")
#user can use plotInfo() function to get index, figure ident and figure map information, then plot figures by index
plotInfo(plots)
#listPlot(object = plots, fig.id = 1)
```


### 2) Heat maps
```{r  results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=6}
# Heatmap for joint clusters
heatMapPlot(object = cbmc, group.by = "jointClusterID", height.rel = 1, adt.label = TRUE)
# Heatmap for RNA clusters
heatMapPlot(object = cbmc, group.by = "rnaClusterID", height.rel = 1, adt.label = TRUE)
# Heatmap for ADT clusters
heatMapPlot(object = cbmc, group.by = "adtClusterID", height.rel = 1, adt.label = TRUE)
```

### 3) RNA and ADT expression
```{r results='hide', message=FALSE, warning=FALSE, echo=TRUE, dpi=300,fig.width=12, fig.height=12}
VlnPlot(cbmc, features = c("rna_CD8A", "adt_CD8", "rna_NCAM1", "adt_CD56", "rna_CD3G", "adt_CD3"), group.by = 'jointClusterID', pt.size = 0, ncol = 2)
```


